Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
fixed one comment
example

\usepackage{pifont}
\newcommand{\lstbg}[3][0pt]{{\fboxsep#1\colorbox{#2}{\strut #3}}}
\lstset{ %
  language=C,
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygray}\bfseries,    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
  frame=single,
  numbers=left,
  stepnumber=1,
  xleftmargin=2em,
  escapeinside={/*!}{!*/},
  moredelim=**[l][\color{mygreen}]{+\ },
  moredelim=*[l][\color{red}]{-\ }
}

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by OpenSSL developers to the
\\CVE-2014-3506 vulnerability},label={lst:vuln}]

+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list) /*!\annotation{lst:func2}!*/
+     return s->max_cert_list;
+  return max_len;
+ }

@@

In the \texttt{dtls1\_reassemble\_fragment} function the value of
\texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the maximum
handshake message size, but \texttt{msg\_len} is allocated to the fragment
buffer in line $37$. Thus, if the fragment was within the allowed size, the
pending handshake message could consume $18MB$ ($16MB$+$2MB$). According to the
message commit, $10$ handshake messages are allowed, so the attacker could
consume approximately $180MB$ per DTLS connection.

@@

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was
inserted to limit the number of bytes allowed in a DTLS handshake message for
\texttt{s}. This function replaces lines $18$---$21$.

~\ref{lst:func2} Checks if the maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise it returns the maximum length.

@@

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a cross-site scripting vulnerability with simple function to sanitize data. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset~\cite{Reis:2017:IJSSE}.
Maintainability is measured separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[caption={Fix provided by nextcloud/server developers to a \\Cross-Site Scripting vulnerability},label={lst:fix},language=php]
 <ul>
    <li class='error'>
       <?php echo $l->t( 'Cloud not found' ); ?><br/>
-      <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+      <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?></p>
    </li>
 </ul>

>>>>

\usepackage{pifont}
\newcommand{\lstbg}[3][0pt]{{\fboxsep#1\colorbox{#2}{\strut #3}}}

\lstdefinestyle{CStyle} {
    language=C,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

\lstdefinestyle{PHPStyle} {
    language=PHP,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

@@

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
\\CVE-2014-3506 vulnerability},label={lst:vuln}]
+ static unsigned long dtls1_max_handshake_message_len(const SSL *s){ /*!\annotation{lst:func1}!*/
+  unsigned long max_len = DTLS1_HM_HEADER_LENGTH + SSL3_RT_MAX_ENCRYPTED_LENGTH;
+  if(max_len < (unsigned long)s->max_cert_list) 
+     return s->max_cert_list;
+  return max_len;
+ }

@@

In the example (Listing~\ref{lst:vuln}), all functions receive as input the \textbf{handshake} represented by \texttt{s}. Every SSL/TLS connection begins with a handshake which is reponsible by the negotiation between two parties. The handshake determines what cryptographic algorithms will be used to encrypt communications, verify the server and establish a secure connection before beginning data transfer. In the \texttt{dtls1\_reassemble\_fragment} function the value of
\texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the maximum
handshake message size, but \texttt{msg\_len} is allocated to the fragment
buffer in line $37$. Thus, if the fragment was within the allowed size, the
pending handshake message could consume $18MB$ ($16MB$+$2MB$). According to the
message commit, $10$ handshake messages are allowed, so the attacker could
consume until $180MB$ per DTLS connection.

@@

~\ref{lst:func1} \texttt{dtls1\_max\_handshake\_message\_len} function was inserted to limit the number of bytes allowed in a DTLS handshake message,
\texttt{s}. Checks if the maximum certificate list size
(\texttt{s-$>$max\_cert\_list}) requires more length than the maximum length
(\texttt{max\_len}). If that condition holds, the function returns the size of
the certificate list, otherwise it returns the maximum length. This function replaces lines $18$-$21$.

@@

There are security flaws that are more difficult to refactor than others. For
instance, implementing secure authentication is not as easy as fixing a cross-site scripting vulnerability with simple function to sanitize data\footnote{More information about \texttt{htmlentities()} at http://php.net/manual/en/function.htmlentities.php (Accessed on \today{})}. Knowing which patterns are more likely to increase maintainability
issues is one step forward to bring awareness to security engineers of what
patterns need more attention. The taxonomy of security patterns used to answer
this question is the same as the one provided by the authors of the dataset~\cite{Reis:2017:IJSSE}.
Maintainability is measured separately for each pattern.

\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix},language=php]
 <ul>
    <li class='error'>
       <?php echo $l->t('Cloud not found');?><br/>
-      <p class='hint'><?php if(isset($_['file'])) echo $_['file']?></p>
+      <p class='hint'><?php if(isset($_['file'])) echo htmlentities($_['file'])?></p>
    </li>
 </ul>
